Online (Bandit) Policy Selection for EASY-Backfilling

### 2017-04-02 paper/src/paper.tex
paper update

Several studies propose to decay $\epsilon$ over time, the exploration being
less important once the estimates for the different policies are reliable
\cite{Tokic:2010}. This strategy has however not been beneficial in our case.
We believe this is due to the properties of the traces we use, as
explained in Section~\ref{sec:experiments}.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{jsm}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. This variability can be so extreme
that even a timeframe of a fews years can be insufficient to compare scheduling
algorithms. In our experience inverting some months in the trace at random can
lead to reversal in the results. This motivated decades of research in workload
modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} proposes to
resample data directly from the original traces. In this work, we use a simple
implementation of these ideas.  More precisely, we want to sample the job
submission process, which is achieved in the following manner: For evey week in
a resampled trace, we iterate over every existing user in the original trace.
For each of these users, we choose a week of the original trace uniformly at
random. The jobs from this user in this original week are chosen to be present
in the week being resampled.  This simple mechanism enables us to preserve the
dependency within a week.  Although the system state may not be independent
from one week to an other, we rely on existing work~\ref{jsspp17} and assume
for these experiments that the system is in a stationary state. While we
do provide guidelines as how to incorporate tracking variant of our algorithms
in~\ref{sec:online} by introducing a decaying discount factor $\lambda$,
this will not be studied in the experiments from this paper, and is left for
future research.

@@

Figures~\ref{fig:small} and~\ref{fig:follow} report the main experimental
result concerning the strategies proposed in this paper. These figures show the
average cumulative improvement of over EASY(FCFS,40h) of the three strategies
we propose in the same format as Figure~\ref{fig:all}.  There are four
strategies displayed on the graph:

\begin{itemize}
  \item Full (simulated) Feedback ($\Delta=$1 day)
  \item Noisy (simulated) Feedback ($\Delta=$1 day)
  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 day, $\epsilon=0.5$)
  \item Random $P \in P_i, 1 \le k \le 100$. ($\Delta=$1 day)
\end{itemize}

@@

\paragraph
\textbf{FCFS vs Random.} Note that randomly taking a policy in the search
set defined in ~\ref{subsec:policies} gives already a much better performance
than EASY(FCFS). This is due

\paragraph
\textbf{Simulated Feedback finds the best policy}
\paragraph
\textbf{Noisy simulations are useful.}
\paragraph
\textbf{Full vs Bandit feedback}

Table~\ref{tab:strat} gives the performance of all variants for $\Delta=$ 1
day and $\Delta=$ 1 week.

Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.

@@

Figure~\ref{fig:mosb} gives insight regarding the behavior of the
Noisy (simulated) Feedback policy. This figure reports the share of the policies chosen
by the strategy on average as a function of time. This is an area plot, which means
that the height of the colored region represents the value. Moreover, the proportion of
times each policies are chosen by the strategy is aggregated across resampled traces.

Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.

@@

Figure~\ref{fig:mosn} is the analogue of Figure~\ref{fig:mosb}, this time for
the Epsilon-Greedy policy.

Observe that \textcolor{orange}{TODO lli  p sumlip sumlip llll sumli psumlips umlip sumil ip sumplip sumsl  ipsumul   ipsumm}.

>>>>

Several studies propose to decay $\epsilon$ over time, the exploration being
less important once the estimates for the different policies are reliable
\cite{Tokic:2010}. This strategy has however not been beneficial in our case.
We believe this is due to the properties of the traces we use, as explained in
Section~\ref{sec:experiments}. The $\epsilon$-greedy strategy is known to be
outperformed in terms of regret by other classical algorithms such as the UCB
or EXP family of algorithms. See~\cite{bubnow} for a precise definition and
thourough treatment of this subject. These algorithms however all rely on
pre-normalized rewards. Here, the scale of $w_{\epsilon}(t;P_i)$ is not known
in advance. While the usual doubling tricks can be devised to go around
sequential issues of scale with good asymptotical performance, they ruin
performance in practice. Accordingly, we only perform experiments using the
simple $\epsilon$-greedy method.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{jsm}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. This variability can be so extreme
that even a timeframe of a fews years can be insufficient to compare scheduling
algorithms. In our experience inverting some months in the trace at random can
lead to reversal in the results. This motivated decades of research in workload
modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} proposes to
resample data directly from the original traces. In this work, we use a simple
implementation of these ideas.  More precisely, we want to sample the job
submission process, which is achieved in the following manner: For evey week in
a resampled trace, we iterate over every existing user in the original trace.
For each of these users, we choose a week of the original trace uniformly at
random. The jobs from this user in this original week are chosen to be present
in the week being resampled.  This simple mechanism enables us to preserve the
dependency within a week.  Although the system state may not be independent
from one week to an other, we rely on existing work~\cite{jsspp17} and assume
for these experiments that the system is in a stationary state. While we
do provide guidelines as how to incorporate tracking variant of our algorithms
in~\ref{sec:online} by introducing a decaying discount factor $\lambda$,
this will not be studied in the experiments from this paper, and is left for
future research.

@@

Figures~\ref{fig:small} and~\ref{fig:follow} report the main experimental
result concerning the strategies proposed in this paper. These figures show the
average cumulative improvement of over EASY(FCFS,40h) of the three strategies
we propose in the same format as Figure~\ref{fig:all}. There are five
strategies displayed on the graph:

\begin{itemize}
  \item Random $P \in P_i, 1 \le k \le 100$. ($\Delta=$1 week)
  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 week)
  \item Epsilon-Greedy (Bandit Feedback) ($\Delta=$1 day)
  \item Full (simulated) Feedback ($\Delta=$1 week)
  \item Noisy (simulated) Feedback ($\Delta=$1 week)
\end{itemize}

@@

\textbf{FCFS vs Random.} Randomly taking a policy in the search set defined in
~\ref{subsec:policies} is better than using the FCFS order. This is due to the
known fact that the FCFS order is not optimal for optimizing the average
waiting time objective, and confirms that the search set is well chosen.  The
value of the period length $\Delta$ has an impact on the performance of the
random policy. Indeed, the shorter period leads to better performance. We
suggest that this is due to the properties of the system under study. Indeed,
when a bad policy is used, the job queue builds up. By switching more
frequently from bad to good policies, the system never has time to build up the
job queue. In other words, while bad decisions impact the system through an
accumulation of waiting jobs, there is less contribution to system inertia from
good decisions, as the number of jobs in queue is below bounded by zero. This
interpretation also means that randomized bandit strategies are robust by
design, and needs more subsequent experimenting to be validated.

\textbf{The Simulated Feedback strategy consistently performs as well as the
best fixed policy.} The values attained by the simulation-based strategy match
the highlighted values from Table~\ref{tab:full}. This corresponds to the curve
labeled "Full" in Figures~\ref{fig:small} and~\ref{fig:follow}. Moreover,
these average cumulative waiting time curves exhibit linearity (as opposed to
being concave), which further shows that the choice of the best policy can be
made rapidly.  Note that the value of $\Delta$ has no impact over the
performance of this strategy.

\textbf{Noisy simulations can be used.} The performance attained by the noisy
variant almost matches that of the simulation-based strategy. This indicates
that it is not necessary to have access to a precise simulator. Rather, we may
hope that an imprecise but unbiased reporting of the performance is enough to
rapidly select a reordering policy .

\textbf{Full vs Bandit feedback.} As expected, the bandit strategy evolves
between the random strategy and the full feedback strategy.
It is able to recover between \textcolor{orange}{X\%} to \textcolor{orange}{X\%}
of the Random-FullFeedback gap.

@@

Figure~\ref{fig:mosb} gives insight regarding the behavior of the Noisy
(simulated) Feedback policy with $\Delta=$ 1 week. This figure reports the
share of the policies chosen by the strategy on average as a function of time.
This is an area plot, which means that the height of the colored region
represents the value. Moreover, the proportion of times each policies are
chosen by the strategy is aggregated across resampled traces.

While this policy obtains an excellent performance, the convergence to a fixed
policy is slow.  This is not harmful \textit{per se} and just means that the
variability in the waiting times is indeed high. Additionally, we observe that
the increasingly dominant policy, LCFS, is the second best overall policy. This
is perhaps due to the estimation bias stemming from the boundary conditions of
the weekly simulations.

@@

Figure~\ref{fig:mosn} is the analogue of Figure~\ref{fig:mosb}, this time for
the Epsilon-Greedy policy with $\Delta$ = 1 week.

Observe that this time the choice is much less stable. The convergence is
extremely slow, which is expected given the variability of the reward we give
to the bandit algorithm.
